A Framework for Modelling Thin Objects: \\ from Meshing to Shell Finite Element Method

### 2010-03-09 MICCAI2010/paper.tex
intro + figures stent
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@85 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\usepackage[english]{babel}
%\usepackage[tight]{subfigure}
%\usepackage{floatrow}
\usepackage{graphicx}
%\usepackage{amsmath}
%\usepackage{multirow}
%\renewcommand*\arraystretch{1.25} % extra vertical space in cells

@@

\section{Introduction}
Cadavers and live animal surgery are often used in the training of surgeons. But ethics concerns and high costs are progressively leading to the elimination of these methods in favor of more modern and effective alternatives. Medical simulators offer an efficient teaching method for physicians. Numerous surgical interventions involve manipulation and deformation of thin anatomical structures and hollow organs. Examples include the interaction between the catheter and the blood vessels during interventional radiology, the deformation of the colon in colonoscopy or the interaction with a beating heart. Simulators of those interventions must showcase a physically realistic behaviour of thin structures to be accepted by the medical community. This is even crucial in the case of colonoscopy for instance as the practitioner only guides himself through the force feedback that he receives from the interaction between the colonoscope and the colon. Although volumic organs are quite accurately simulated in recent simulators hollow organs are poorly modelled, which limits the type of medical procedures that a simulator can provide.

We aim to fill this gap by offering a framework able to handle all the steps from meshing to realistic simulation of hollow organs. Not only we propose an efficient yet simple co-rotational finite element formulation for triangular shells but we also present a method to generate a mesh of any hollow organ with shell-type curved elements in an attempt to optimise even more the computation of deformations by using fewer elements. % Finally we show the pertinence of our approach by coupling tetrahedral and shell elements in an advanced liver model.

\section{A physical model for thin structures}

\subsection{Related work}

@@

\begin{equation}
\textbf{K}_e = \int_v \textbf{J} \boldsymbol\chi \textbf{J}^{T} dV
\end{equation}
where $\textbf{J}$ is the strain-displacement matrix and $\boldsymbol\chi$ embodies the material's behaviour. The stiffness matrix in the global frame is eventually obtained using the rotation matrix of the element: $\textbf{K} = \textbf{R}^{T} \textbf{K}_e \textbf{R} $ where $\textbf{R}$ describes the rotation of the (triangular) element with respect to its initial configuration.

@@

\section{Meshing a curved surface}

\subsection{Related work}

@@

\section{Results}
\subsection{Meshing of anatomical structures (error maps)}

\subsection{Deformations}
\subsubsection{Computation times on 3 examples}
\subsubsection{Coupling between tetrahedra and shells for more advanced modelling}

>>>>

\usepackage[english]{babel}
%\usepackage[tight]{subfigure}
%\usepackage{floatrow}
\usepackage{graphicx}
%\usepackage{amsmath}
%\usepackage{multirow}
%\renewcommand*\arraystretch{1.25} % extra vertical space in cells
\usepackage{amsmath}

@@

\section{Introduction}

Human body is composed of various deformable anatomical structures. 
Realistic modeling of organ deformations is a challenging research field that leads the way to numerous clinical applications including: medical training and rehearsal systems, patient-specific planning of surgical procedure and per-operative guidance based on simulation.
In all these cases, the clinician needs fast update of the deformation model to obtain a real-time display of the computed deformations.  
To realize this kind of interactive systems, one face the trade-off problem between accuracy of the model and the need of fast computations.

An other key challenge of real-time human soft-tissue modeling is the variousness of the mechanical behaviors. 
It seems unrealistic to target a unique model for all tissues even if there are obvious advantages to make maximum use of optimized generic models.
Yet most of previous work focus on volume models that are able to capture the behavior of volume organs like liver \cite{} or brain tissues \cite{}.
In contrast, this paper seeks to propose a solution for simulating, in real-time, deformation of thin anatomical structures, whose volume is negligible compared to their surface area.
Examples of such anatomical structures include hollow structures, like the wall of blood vessels or the colon but also the envelope of volume organs, like the capsule of Glisson.

In the following of this paper, we present a framework for realistic and real-time simulation of thin organs.
Our method, presented in section \ref{sec:model} is based on shell theory and computes the internal forces using Finite Element Modeling (FEM) on curved triangles.
The design of the elements relies on a polynomial interpolation that is also use to capture the interactions with other models, including volume models.
We also present an automatic process to obtain meshes from image based reconstruction in an attempt to optimize  the computation using fewer elements  (see section \ref{sec:mesh}).
The benefits of our approach (meshing of a curved surface, fast computation and possible interactions with volume models) are illustrated using various examples presented in section \ref{sec:results}.

\section{A physical model for thin structures}
\label{sec:model}
\subsection{Related work}

@@

\begin{equation}
\textbf{K}_e = \int_v \textbf{J} \boldsymbol{\chi} \textbf{J}^{T} dV
\end{equation}
where $\textbf{J}$ is the strain-displacement matrix and $\boldsymbol\chi$ embodies the material's behaviour. The stiffness matrix in the global frame is eventually obtained using the rotation matrix of the element: $\textbf{K} = \textbf{R}^{T} \textbf{K}_e \textbf{R} $ where $\textbf{R}$ describes the rotation of the (triangular) element with respect to its initial configuration.

@@

\section{Meshing a curved surface}
\label{sec:mesh}
\subsection{Related work}

@@

\section{Results}
\label{sec:results}
\subsection{Meshing of anatomical structures (error maps)}

\subsection{Deformations}
\subsubsection{Computation times on 3 examples}
\subsubsection{Coupling between tetrahedra and shells for more advanced modelling}
\begin{figure}
\centering
\includegraphics[height=4cm]{images/stent_begin.png}
\includegraphics[height=4cm]{images/stent_end.png}
\caption {Avant... Apres...}
\label{fig-stent}
\end{figure}
